---
title: 'PRELIMINARY DATA: Redd Dewatering Estimates for Keswick Fall Flow Scenarios'
author: "BDO Science Division"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document: default
  word_document: default
  html_document: default
always_allow_html: true
---

```{r setup, echo=FALSE, include = FALSE}
#loading libraries and pulling in relevant data from excel files
library(tinytex)
library(data.table)
library(plyr)
library(ggplot2)
library(tidyverse)
library(readxl)
library(zoo)
library(kableExtra)
library(janitor)
library(lubridate)

```

This script constructs real-time winter-run redd dewatering estimates based on most recent data available from CDFW (`r updatedReddInfoDate`) for winter-run data and dewatering estimates from USFWS (2006; see citation). Data are also available in `r yr` Winter-run Data file.xls online at [calfish.org](https://gcc02.safelinks.protection.outlook.com/?url=https%3A%2F%2Fwww.calfish.org%2FProgramsData%2FConservationandManagement%2FCentralValleyMonitoring%2FCDFWUpperSacRiverBasinSalmonidMonitoring.aspx&data=05%7C01%7Clelliott%40usbr.gov%7C689ebb9a6c8243b4f96c08da90f5c542%7C0693b5ba4b184d7b9341f32f400a5494%7C0%7C0%7C637981682646098788%7CUnknown%7CTWFpbGZsb3d8eyJWIjoiMC4wLjAwMDAiLCJQIjoiV2luMzIiLCJBTiI6Ik1haWwiLCJXVCI6Mn0%3D%7C3000%7C%7C%7C&sdata=A1eQkWPxbkXxnzEvc2K8%2FTmslZ8H8zvxdks3%2F78Yrvw%3D&reserved=0).

This document is also now available on [SacPAS](https://www.cbr.washington.edu/sacramento/workgroups/usst.html#redd_dewater).  However, data on the SacPAS webpage may not immediately reflect data in the document as updates to the webpage with newest flow and redd data may be pending.

Please note that all data are preliminary until data collection is finalized. Likewise, there are uncertainties with forecasts which may lead to changes in proposed operations.

# Current Winter-run Chinook Salmon Redd Count

As of `r countDate`, the unexpanded redd count is **`r reddCount`** Winter-run redds. It is important to note that until data collection is completed for the year these are the **minimum** number of possible redds. The Winter-run number will always expand upon final analysis but gives an in-season guard rail of the minimum number of redds this year.

Given that the number of Winter-run redds is always larger than the early season carcass counts, an expansion number based on historic data is multiplied by the carcass count to estimate the total number of redds for the season before the end of the season's final estimate is developed and the final redd count is known. Average 2005-2022 expansion was 1.98 * the total redd count, and thus we focus on an expansion factor of 2 to represent expected final redd count and support decision-making.

```{r, echo = FALSE, message=FALSE}
# create a table for population expansions

if (knitr::is_html_output()) {
  knitr::kable(pop_tab,
               caption = 'Estimated total number of Winter-run redds and resulting number of redds that represent 1% of the population. Estimated total redds are based on current count and expansion number representing average 2005-2022 expansion.')
} else if (knitr::is_latex_output()) {
  knitr::kable(pop_tab,
               caption = 'Estimated total number of Winter-run redds and resulting number of redds that represent 1% of the population. Estimated total redds are based on current count and expansion number representing average 2005-2022 expansion.')
} else if (knitr::pandoc_to("docx")) {
  flextable::flextable(pop_tab) %>%
    flextable::set_caption(caption = 'Table 1. Estimated total number of Winter-run redds and resulting number of redds that represent 1% of the population. Estimated total redds are based on current count and expansion number representing average 2005-2022 expansion.') %>%
    flextable::width(width = 1.5, unit = "in") %>%
    flextable::font(font = 'Cambria', part = 'all')
}
```

## Chinook Salmon Dewatered Redd Estimates

As of `r format(updatedReddInfoDate, '%B %d, %Y')`, **`r nrow(subset(reddsAll, reddsAll$Status == "EMERGED"))`** Winter-run redds have **emerged** and **`r nrow(subset(reddsAll, reddsAll$Status == "DEWATERED"))`** have been **dewatered**. This leaves **`r nrow(subset(reddsAll, reddsAll$Status == "OK"))`** shallow water redds of concern.

There is no real time data on fall-run redd counts. Estimates are predicted based on estimated dewatering percentages from USFWS (2006) and spring-run and fall-run spawn timing based on fresh female carcasses encountered by week from 2003 through 2023. Emergence timing were predicted from water temperatures below Keswick in 2018 which most closely aligns with 2024 operations targeting 53.5 F at Clear Creek. Fall-run dewatered redd estimates range from **`r min(summary[2])`** to **`r max(summary[2])`%**.  Note that fall-run dewatering estimates are likely overestimated using the dewatering percentages from USFWS (2006), and likely do not reflect actual dewatering percentages and should only be used for comparative purposes between scenarios.  A comparative analysis between field and modeled dewatering percentages by Gosselin and Beer (2024) can be found here: https://www.cbr.washington.edu/sacramento/fishmodel/Note_on_Redd_Dewatering_Observed_v_Predicted.pdf.

## Carryover Effects to Next Year Winter-run Brood  
```{r, echo=FALSE, message = FALSE}
#do any EOS fall below 2200 TAF?  Only runs after September 1 or more scenarios are proposed
if (Sys.Date() >= as.Date(paste0(yr, "-09-01"))) {
  
  EOS <- 2717
  eosList <- list()
  
  eosList <- eosList %>% append(
    list(total_volume_as = (colSums(filter(kesFlowReal2, date >= paste0(yr,'-08-01') & date < paste0(yr,'-10-01'))[, -1])*1.983/1000)),
    after = length(eosList)
  )

  filtered_data <- filter(kesAug90, date >= paste0(yr,'-08-01') & date < paste0(yr,'-10-01'))
  eos_aug90 <- colSums(filtered_data[, -1]) * 1.983 / 1000

  eosList <- eosList %>% append(
    list(Difference_from_aug90 = (eos_aug90 - colSums(filter(kesFlowReal2, date >= paste0(yr,'-08-01') & date < paste0(yr,'-10-01'))[, -1])*1.983/1000)),
    after = length(eosList)
  )

  eosList <- eosList %>% append(
    list(Resulting_EOS = (EOS - (eos_aug90 - colSums(filter(kesFlowReal2, date >= paste0(yr,'-08-01') & date < paste0(yr,'-10-01'))[, -1])*1.983/1000))),
    after = length(eosList)
  )

  if (min(eosList$Resulting_EOS) > 2200) {
    TDM <- "All proposed scenarios are anticipated to have EOS storage greater than the 2200 TAF threshold and therefore would not be expected to contribute to TDM impacts to winter-run chinook salmon in the subsequent year"
  } else {
    TDM <- "One or more proposed scenarios are anticipated to have EOS storage below the 2200 TAF threshold and therefore would be expected to contribute to TDM in the subsequent year, though scenarios above the threshold are not expected to contribute to TDM impacts to winter-run chinook salmon in the subsequent year"
  }

}

```

An analysis on the relationship between winter-run chinook salmon temperature dependent mortality relationship and Shasta Reservoir end-of-year storage suggests a threshold of 2,200 TAF end of September Shasta Storage to assess the impacts of TDM impacts on next year’s cohort. Next year’s cohort is expected to experience minimal TDM impacts when end of September Shasta Storage is greater than this threshold, while values lower than 2,200 TAF are correlated with potentially more negative TDM impacts.  As of August, End of September Shasta Storage is expected to be **`r EOS`** TAF.`r TDM` (see Table 2).

## Preliminary Predicted Results  
```{r, echo=FALSE, message = FALSE}
if (knitr::is_html_output()) {
  knitr::kable(summary_table,
               caption = paste0('Summary of water volume and winter-run and fall-run dewatering estimates related to flow scenarios. Each scenario uses actual flow-to-date as of most current report and proposed flows for the remainder of the incubation period. Percentage of the population lost is based on the ', countDate, ' count of ', reddCount, ' Winter-run redds. See Scenario Descriptions file for additional information on each scenario.')
               )
} else if (knitr::is_latex_output()) {
  knitr::kable(summary_table,
               caption = paste0('Summary of water volume and winter-run and fall-run dewatering estimates related to flow scenarios. Each scenario uses actual flow-to-date as of most current report and proposed flows for the remainder of the incubation period. Percentage of the population lost is based on the ', countDate, ' count of ', reddCount, ' Winter-run redds. See Scenario Descriptions file for additional information on each scenario.')
               ) %>%
    kable_styling(font_size = 8) %>%
    column_spec(1, width = "4cm")
} else if (knitr::pandoc_to("docx")) {
  flextable::flextable(summary_table) %>%
    flextable::set_caption(caption = paste0('Table 2. Summary of water volume and winter-run and fall-run dewatering estimates related to flow scenarios. Each scenario uses actual flow-to-date as of most current report and proposed flows for the remainder of the incubation period. Percentage of the population lost is based on the ', countDate, ' count of ', reddCount, ' Winter-run redds. See Scenario Descriptions file for additional information on each scenario.')) %>%
    flextable::width(width = 1, unit = "in") %>%
    flextable::align(j = -1, align = 'center', part = 'all') %>%
    flextable::font(font = 'Cambria', part = 'all')
}
```



```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.cap = 'Actual or estimated emergence dates of SRWC redds and actual or estimated dewatering flow for the September-October estimated redd emergence dates as compared to Keswick flow (in cfs) of proposed management alternatives. Points represent dewatered (De), emerged (Em), or remaining (Re) redds. Numbers inside of points indicate how many redds share that estimated emergence date and actual/estimated dewatering flow. Points that fall above/to the right of a flow alternative line are expected to be dewatered given that management alternative is followed. Points that fall below/to the left of/on a flow alternative line are not expected to be dewatered, given that management alternative is followed. Shaded gray box shows period of real-time flow data; dashed black line equals KWK gauge flow and solid black line equals KES flow (from [SacPas](https://www.cbr.washington.edu/sacramento/data/query_river_table.html)).', fig.dim = c(8,5)}

#, fig.dim = c(8,5)
reddsDat <- reddsAll
reddsTable <- setDT(reddsDat)[,list(Count=.N), names(reddsDat)]
ymin <- round_any(min(reddsTable$ACTUAL.or.ESTIMATED..DEWATER.FLOW), 1000, f = floor)
#ymax <- round_any(max(reddsTable$ACTUAL.or.ESTIMATED..DEWATER.FLOW), 250, f = ceiling)
ymax <- round_any(max(allflows$Flow)+250, 500, f = ceiling)
todays_date <- Sys.Date()
mid <- as.Date(paste0(yr,'-08-01')) + floor((todays_date - as.Date(paste0(yr,'-08-01')))/2)
for(r in 1:nrow(reddsTable)){
  if(reddsTable$Status[r] == "EMERGED"){
    reddsTable$cols[r] <- "Emerged"
  } else {
    if(reddsTable$Status[r] == "DEWATERED"){
      reddsTable$cols[r] <- "Dewatered"
    } else {
      reddsTable$cols[r] <- "Remaining"
    }
  }
}
reddsTable <- reddsTable %>% 
  rename('Date' = 'Estimated.Date.of.Emergence', 'Flow' = 'ACTUAL.or.ESTIMATED..DEWATER.FLOW') %>% 
  mutate(cols = factor(cols, levels = c('Remaining', 'Emerged', 'Dewatered'), labels = c('Re','Em', 'De')))

flows <- kesFlowReal %>% 
  gather(key = Alts, value = Flow, contains('cfs')) %>%
  filter(date >= todays_date) %>%
  mutate(Alts = gsub('_cfs','',Alts))

ggplot() + geom_line(flows, mapping = aes(x = date, y = Flow, color = Alts), size = 0.75) +
  geom_line(allflows, mapping = aes(x = Date, y = Flow, linetype = Gage), size = 0.75) +
  scale_linetype_discrete(guide = "none") +
  geom_label(reddsTable, mapping = aes(x = Date, y = Flow, 
                                       label = Count, fill = cols), min.segment.length = 4,
             postion = 'dodge', size = 3, label.padding = unit(0.25, 'lines'),label.r = unit(.35, 'lines'), 
             fontface = 'bold', alpha = 0.6) +
  ylim(ymin, ymax) +
  xlim(min(as.Date(paste0(yr,'-08-01'))), (max(reddsTable$Date) + 5)) +
  labs(x = 'Date', y = 'Flow (cfs)', fill = 'Redd Status') +
  theme(legend.position = 'bottom', 
        legend.text = element_text(size = 12), 
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 12),
        legend.title = element_text(size = 14),
        legend.key.size = unit(0.3, 'cm')) +
  scale_fill_manual(values = c('lightgrey', 'steelblue3', 'darkorange')) +
  annotate(geom = 'rect', xmin = as.Date(paste0(yr,'-08-01')), xmax = max(todays_date), 
           ymin = 3000, ymax = max(allflows$Flow) + 100, fill = 'darkgrey', color = 'black', alpha = 0.2, linetype = 'dotted') +
  annotate(geom = 'text', x = (mid + 0.5), y = max(allflows$Flow) + 400, size = 3.5, 
           fontface = 'italic', label = 'Actual Flows') +
  annotate(geom = 'text', x = as.Date('2023-09-03'), y = 9350, size = 3, label = 'KWK') +
  annotate(geom = 'text', x = as.Date('2023-09-03'), y = 7750, size = 3, label = 'KES')

```


```{r echo=FALSE, message=FALSE}
worksheetSD <- read_xlsx(paste0('External_data/FlowScen/',MaxFile), sheet = "Scenario Description",range="A1:A1000",col_names=F,col_types="text")
# remove extraneous lines, keeping lines with "=" in it
keep=apply(worksheetSD,2,function(x.in) str_detect(x.in,"="))
worksheetSD=worksheetSD[keep,]
worksheetSD=worksheetSD[!is.na(worksheetSD[,1]),]
# split the Alts from the descriptions
num.in = dim(worksheetSD)[1]
scenarioDesc=data.frame(matrix(nrow=num.in,ncol=2))
colnames(scenarioDesc)=c("Scenario","Description")
# loop through dataframe, split original dataframe column 1 on literal "=" characters and trim leading and trailing space, create new dataframe with 2 columns "Scenario" and "Description"
# populates scenarioDesc dataframe
for(i in 1:num.in){
	scenarioDesc[i,]=str_trim(str_split(worksheetSD[i,1],"=")[[1]])
}

scenarioDesc <- scenarioDesc %>%
  mutate(Key = gsub("[^[:alnum:]]+", "", Scenario)) %>%
  mutate(Key = tolower(Key)) %>%
  filter(Key %in% scen) %>%
  mutate(Scenario = paste0(Scenario,' (',Key,')')) %>%
  select(-Key)
#use scenarios defined in first chunk of code to filter the scenarioDesc dataframe



#filter based on scenario name
#scenarioDesc <- scenarioDesc %>% 
  #filter(grepl(paste(toupper(scenarios2), collapse ='|'), toupper(Scenario)))
if (knitr::is_html_output()) {
  knitr::kable(scenarioDesc, 
               caption = 'Description of scenarios being considered and compared by the Upper Sacramento Scheduling Team. Scenario name includes the shorthand notion in parentheses for cross-referencing with graph and tables.')
} else if (knitr::is_latex_output()) {
  knitr::kable(scenarioDesc, 
               caption = 'Description of scenarios being considered and compared by the Upper Sacramento Scheduling Team. Scenario name includes the shorthand notion in parentheses for cross-referencing with graph and tables.') %>%
    column_spec(1, width = "3cm") %>%
    column_spec(2, width = "13cm")
} else if (knitr::pandoc_to("docx")) {
  flextable::flextable(scenarioDesc) %>%
    flextable::set_caption(caption = 'Table 3. Description of scenarios being considered and compared by the Upper Sacramento Scheduling Team.  Scenario name includes the shorthand notion in parentheses for cross-referencing with graph and tables.') %>%
    flextable::width(1, width = 1.75, unit = "in") %>%
    flextable::width(2, width = 5, unit = "in") %>%
    flextable::font(font = 'Cambria', part = 'all')
}
```
# References

Gard, Mark. 2006. Relationships between flow fluctuations and redd dewatering and juvenile stranding for Chinook Salmon and Steelhead in the Sacramento River between Keswick Dam and Battle Creek. 94 pages.

Gosselin, J.L. and W.N. Beer. 2024. Sacramento River Winter-run Chinook Salmon Redd Dewatering: a Note on Comparing Observed and Predicted. Central Valley Prediction and Assessment of Salmon (SacPas; https://www.cbr.washington.edu/sacramento/). Columbia Basin Research, School of Aquatic and Fishery Sciences, University of Washington.
