---
title: "PRELIMINARY DATA: Redd Dewatering Estimates for Keswick Fall Flow Scenarios"
author: "BDO Science Division"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r, echo=FALSE, include = FALSE}
reddCount <- 347
countDate <- "August 13, 2023"
updatedReddInfoDate <- "September 6, 2023"
fallreddCount<- 1000

scenarios <- c("Alt.1b", "Alt.3c", "Alt.3d", "Alt.3f", "Alt.3g")

#KWK daily flows url.  This url will have to be updated.
library(rvest)
kwk_webpage <- read_html('https://www.cbr.washington.edu/sacramento/tmp/riverdaily_1694103865_328.html')
kes_webpage <- read_html('https://www.cbr.washington.edu/sacramento/tmp/riverdaily_1694103884_126.html')
#show_extra = TRUE
```

This script constructs real-time winter-run redd dewatering estimates based on most recent data available from CDFW (`r updatedReddInfoDate`) for winter-run data and dewatering estimates from USFWS (2006; see citation). Data are also available in the YYYY Winter-run Data file.xls online at [calfish.org](https://gcc02.safelinks.protection.outlook.com/?url=https%3A%2F%2Fwww.calfish.org%2FProgramsData%2FConservationandManagement%2FCentralValleyMonitoring%2FCDFWUpperSacRiverBasinSalmonidMonitoring.aspx&data=05%7C01%7Clelliott%40usbr.gov%7C689ebb9a6c8243b4f96c08da90f5c542%7C0693b5ba4b184d7b9341f32f400a5494%7C0%7C0%7C637981682646098788%7CUnknown%7CTWFpbGZsb3d8eyJWIjoiMC4wLjAwMDAiLCJQIjoiV2luMzIiLCJBTiI6Ik1haWwiLCJXVCI6Mn0%3D%7C3000%7C%7C%7C&sdata=A1eQkWPxbkXxnzEvc2K8%2FTmslZ8H8zvxdks3%2F78Yrvw%3D&reserved=0).


Please note that all data are preliminary until data collection is finalized. Likewise, there are uncertainties with forecasts which may lead to changes in proposed operations. 

```{r setup, echo=FALSE, include = FALSE}
library(tinytex)
library(data.table)
library(plyr)
library(lubridate)
library(ggplot2)
library(ggrepel)
library(dplyr)
library(tidyverse)

#this function will convert the character values (from xlsx format of numbers with commas) to numeric class values
setClass("num.with.commas")
setAs("character", "num.with.commas", 
        function(from) as.numeric(gsub(",", "", from) ) )

#Read in redd data, converting flow to a numeric and date from character to date value
redds <- read.csv("Redds.csv", colClasses=c('character', 'character', 'num.with.commas'))
redds$Estimated.Date.of.Emergence <- as.Date(redds$Estimated.Date.of.Emergence, "%m/%d/%Y")
#redds$Estimated.Date.of.Emergence <- as.Date(redds$Estimated.Date.of.Emergence,  "%Y-%m-%d")
reddsAll <- redds

#Read in KES flow data, converting flow to a numeric and date from character to date value
kesFlow <- read.csv("kesFlow.csv")
kesFlow <- read.csv("kesFlow.csv", colClasses=c(rep('character', length(kesFlow))))
kesFlow$Date <- as.Date(kesFlow$Date, "%m/%d/%Y")
#kesFlow$Date <- as.Date(kesFlow$Date, "%Y-%m-%d")
kesFlow <- kesFlow %>%
  as_tibble %>%
  select(Date, scenarios) %>%
  as.data.frame()

realFlow <- read.csv("realtimeKES.csv")
realFlow$Date <- as.Date(realFlow$Date, "%m/%d/%Y")
actualDate <- max(realFlow$Date)

#import fall-run spawn info
Spawn_info <- read.csv('spawn.csv')
Spawn_info$Date <- as.Date(Spawn_info$Date, format = '%m/%d/%Y')
Spawn_info <- Spawn_info %>% mutate(Emerg_Date = as.Date(Date + Avg_emerg)) #calculate emergence date

#import fall-run model
fall_model <- readRDS('fall_model.rds')

#`r if(show_extra){
#print("this will only print when show.text is TRUE")
#str(redds)
#str(kesFlow)
#str(realFlow)}`

```

```{r, echo = FALSE, include = FALSE}
library(rvest)
library(dplyr)
library(lubridate)


#list all table objects on webpage and print tbls to look for the proper table
kwk_tbls <- html_nodes(kwk_webpage, 'table')
kwk_tbls

#the proper table appears to be 6th object on the tbls list so input that into the code
kwk_flows <- kwk_webpage %>% html_nodes("table") %>% .[6] %>% html_table(fill = TRUE)

#convert to a dataframe and set the first row as the column headings
kwk_flows <- as.data.frame(kwk_flows)
#colnames(kwk_flows) <- kwk_flows[1, ]
colnames(kwk_flows) <- c("Project", "Date", "KWK")
kwk_flows <- kwk_flows[-1,]
kwk_flows$Date <- as.Date(kwk_flows$Date, "%Y-%m-%d")

kwk_flows <- kwk_flows %>%
  select(Date, KWK) %>%
  filter(Date >= as.Date("2023-08-01", "%Y-%m-%d"))





#list all table objects on webpage and print tbls to look for the proper table
kes_tbls <- html_nodes(kes_webpage, 'table')
kes_tbls

#the proper table appears to be 6th object on the tbls list so input that into the code
kes_flows <- kes_webpage %>% html_nodes("table") %>% .[6] %>% html_table(fill = TRUE)

#convert to a dataframe and set the first row as the column headings
kes_flows <- as.data.frame(kes_flows)
colnames(kes_flows) <- kes_flows[1, ]
kes_flows <- kes_flows %>%
  select(Project, contains("Date"), contains("Outflow"))
colnames(kes_flows) <- c("Project", "Date", "KES")
kes_flows <- kes_flows[-1,]
kes_flows$Date <- as.Date(kes_flows$Date, "%Y-%m-%d")

kes_flows <- kes_flows %>%
  select(Date, KES) %>%
  filter(Date >= as.Date("2023-08-01", "%Y-%m-%d"))



#import actual Keswick flows from USGS
#library(dataRetrieval)
#start_date <- as.Date('2023-08-01')
#end_date <- Sys.Date()
#kes_flow <- readNWISdata(sites = '11370500', parameterCd = '00060', startDate = start_date, endDate = end_date)
```


```{r, echo = FALSE}
# Thousand Acre Feet

kesFlowReal <- merge(kesFlow, kes_flows, by = "Date", all.x = TRUE, all.y = TRUE)
for(a in 1:(length(kesFlow)-1)){
 kesFlowReal[ ,  paste0("A", substr(names(kesFlow)[a + 1], 4,6), "cfs")] <- NA
 kesFlowReal[ ,  paste0("A", substr(names(kesFlow)[a + 1], 4,6), "taf")] <- NA
}

for(a in 1:(length(kesFlow)-1)){
  for(r in 1: nrow(kesFlowReal)){
    if(is.na(kesFlowReal$KES[r])){
      kesFlowReal[r, paste0("A", substr(names(kesFlow)[a + 1], 4,6), "cfs")] <- as.numeric(kesFlowReal[r, paste0("Alt", substr(names(kesFlow)[a + 1], 4,6))])
      kesFlowReal[r, paste0("A", substr(names(kesFlow)[a + 1], 4,6), "taf")] <- (as.numeric(kesFlowReal[r, paste0("Alt", substr(names(kesFlow)[a + 1], 4,6))])*1.983)/1000
    } else{
      kesFlowReal[r, paste0("A", substr(names(kesFlow)[a + 1], 4,6), "cfs")] <- as.numeric(kesFlowReal$KES[r])
      kesFlowReal[r, paste0("A", substr(names(kesFlow)[a + 1], 4,6), "taf")] <- (as.numeric(kesFlowReal$KES[r])*1.983)/1000
    }
  }
}

septTAF <- subset(kesFlowReal, months(kesFlowReal$Date) == "September")
octTAF <- subset(kesFlowReal, months(kesFlowReal$Date) == "October")

```

```{r, echo = FALSE}
# September Flow Average
septKesFlow <- subset (kesFlow, kesFlow$Date < '2023-10-01' & kesFlow$Date >= '2023-9-01')
octKesFlow <- subset (kesFlow, kesFlow$Date >= '2023-10-01' & kesFlow$Date < '2023-11-01')
```

```{r, echo = FALSE, message=FALSE}
#estimate fall-run redds dewatered

#transform kesFlowReal dataframe for model
redd_model <- kesFlowReal %>% 
  gather(key = Scenarios, value = Spawn_Flows, contains('cfs')) %>%
  select(Date, Scenarios, Spawn_Flows) %>% 
  mutate(Boards = if_else(Date < as.Date('2023-11-01'), 'in', 'out')) %>% 
  mutate(Spawn_Flows = as.numeric(Spawn_Flows))

#Join the spawning information and add the minimum flow each emergence date will be subjected to
redd_model <- redd_model %>% left_join(Spawn_info, by = 'Date') %>% group_by(Scenarios) %>%
  mutate(Dewater = sapply(Emerg_Date, function(edate) {
    if (!is.na(edate)) {
      min(Spawn_Flows[Date <= edate], na.rm = TRUE)
    } else {
      NA
    }
  })) %>%
  ungroup() %>% na.omit() %>% group_by(Date) %>% mutate(Redds = fallreddCount * Per_spawn) %>% ungroup()

#apply model
pred <- predict.lm(fall_model, newdata = redd_model, type = 'response')
redd_model <- cbind(redd_model, pred) 
redd_model <- pmax(redd_model) #rounds negative predictions up to 0
redd_model <- redd_model %>% mutate(pred = if_else(Dewater >= Spawn_Flows, 0, pred)) #converts equal flows to 0

summary <- redd_model %>% mutate(Redds_lost = pred * Per_spawn) %>% group_by(Scenarios) %>% 
  summarize(Redds_lost = round((sum(Redds_lost)*100), 1)) %>%
  rename('Fall-run dewatered (%)' = 'Redds_lost')
table2 <- setDT(as.data.frame(t(summary)), keep.rownames = TRUE, check.names = TRUE)
names(table2) <- table2 %>% slice(1) %>% unlist()
table2 <- table2 %>% slice(-1)
```

# Current Winter-run Chinook Salmon Redd Count
As of `r countDate`, the unexpanded redd count is **`r reddCount`** Winter-run redds.   It is important to note that until data collection is completed for the year these are the **minimum** number of possible redds.  The Winter-run number will always expand upon final analysis but gives an in-season guard rail of the minimum number of redds this year. From 2018-2021, female expansion has ranged from 0.31 to 1.31 with a 0.7 average, thus we may expect the final number of Winter-run redds to be closer to `r round(reddCount*1.7, 0)` redds using average expansion on data this year, and using this value, `r 0.01*round(reddCount*1.7, 0)` redds dewatered would be at the 1% population loss.


## Chinook Salmon Dewatered Redd Estimates
As of `r updatedReddInfoDate`, **`r nrow(subset(reddsAll, reddsAll$Status == "EMERGED")) `** Winter-run redds have **emerged** and **`r nrow(subset(reddsAll, reddsAll$Status == "DEWATERED"))`** have been **dewatered**. This leaves **`r nrow(subset(reddsAll, reddsAll$Status == "OK"))`** shallow water redds of concern.  

There is no real time data on fall-run redd counts. Estimates are predicted based on estimated dewatering percentages from USFWS (2006) and spring-run and fall-run spawn timing based on fresh female carcasses encountered by week from 2003 through 2022. Estimated emergence dates are based on river water temperatures during fall-run period at the CCR gauge in 2011 that most closely mimics the 2023 summer water temperatures. Fall-run dewatered redd estimates range from **`r min(summary[2]) `** to **`r max(summary[2])`%**.



```{r, echo = FALSE}
# Estimated number of redds dewatered
for(a in 1:(length(kesFlow)-1)){
 redds[ ,  paste0("dewateredAlt", substr(names(kesFlow)[a + 1], 4,6))] <- NA
}

redds <- subset(redds, redds$Status == "OK" | redds$Status == "DEWATERED")

for(r in 1:nrow(redds)){ #loop over redds
  for(f in 1:nrow(kesFlow)){ #loop over KES flow
    for(a in 1:(length(kesFlow)-1)){ #loop over alternatives
      if(kesFlow$Date[f] == redds$Estimated.Date.of.Emergence[r]){ #if dates match
        if(kesFlow[f, paste0("Alt", substr(names(kesFlow)[a + 1], 4,6))] <= redds$ACTUAL.or.ESTIMATED..DEWATER.FLOW[r]){ #If KES flow is < or = redd dewatering flow, 1 (dewatered)
          redds[r, paste0("dewateredAlt", substr(names(kesFlow)[a + 1], 4,6))] <- 1
        } else{ #If KES flow is greater than redd dewatering flow, mark 0 (not dewatered)
          redds[r, paste0("dewateredAlt", substr(names(kesFlow)[a + 1], 4,6))] <- redds[r, paste0("dewateredAlt", substr(names(kesFlow)[a + 1], 4,6))] 
          } #ends else
      }
    } #ends if date is the same
  } #end looping over KES flow dataframe
} #end looping over redds dataframe
#Based on the `r countDate` redd count:
```


```{r, echo = FALSE}
# Estimated number of redds dewatered with a 100cfs buffer
for(a in 1:(length(kesFlow)-1)){
 redds[ ,  paste0("dewateredAlt", substr(names(kesFlow)[a + 1], 4,6), "Buffer")] <- NA
}

for(r in 1:nrow(redds)){ #loop over redds
  for(f in 1:nrow(kesFlow)){ #loop over KES flow
    for(a in 1:(length(kesFlow)-1)){ #loop over alternatives
      if(kesFlow$Date[f] == redds$Estimated.Date.of.Emergence[r]){ #if dates match
        if(as.numeric(kesFlow[f, paste0("Alt", substr(names(kesFlow)[a + 1], 4,6))]) <= redds$ACTUAL.or.ESTIMATED..DEWATER.FLOW[r] + 250){ #If KES flow is < or = redd dewatering flow, 1 (dewatered)
          redds[r, paste0("dewateredAlt", substr(names(kesFlow)[a + 1], 4,6), "Buffer")] <- 1
        } else{ #If KES flow is greater than redd dewatering flow, mark 0 (not dewatered)
          redds[r, paste0("dewateredAlt", substr(names(kesFlow)[a + 1], 4,6), "Buffer")] <- redds[r, paste0("dewateredAlt", substr(names(kesFlow)[a + 1], 4,6))] 
          } #ends else
      }
    } #ends if date is the same
  } #end looping over KES flow dataframe
} #end looping over redds dataframe
#Based on the `r countDate` redd count:

```


# Table
**Table 1.** Average September and October Keswick (KES) Flow in cfs, total water volume of each alternative for August through October and September through February in TAF, estimated numbers of SRWC redds dewatered, and percent of population that would be lost under each of the proposed alternatives. KES Flow data uses actual flow-to-date as of `r max(realFlow$Date)` and proposed flows for the remainder of the incubation period. Redd dewatering is considered at the actual or estimated dewatering flow and with a 250 cfs buffer applied to the actual/estimated dewatering flow. Percentage of the population lost is based on the `r countDate` count of `r reddCount` Winter-run redds and updated redd counts may be available soon. See Scenario Descriptions file for additional information on each scenario. 


```{r, echo=FALSE, results= TRUE}
Metric <- c("Avg Sept Flow (cfs)", "Avg Oct Flow (cfs)", "Sept-Feb Total Volume (TAF)", "Aug-Oct Total Volume (TAF)", "Winter-run Redds dewatered", "Winter-run Percent Lost (based on current count)", "Winter-run Percent Lost (based on expansion of 0.7)", "Winter-run Percent Lost (based on expansion of 0.3)", "Winter-run Redds Dewatered (w/ 250 cfs buffer)", "Winter-run Percent Lost (w/ 250 cfs buffer)")
table <- data.frame(Metric)


d <- rep(NA, length(kesFlow)-1)
dflist = vector("list", length(kesFlow)-1) # creating an empty list
dfnames <- rep(NA, length(kesFlow)-1)

for(a in 1:(length(kesFlow)-1)){
 d[a] = paste0("Alt", substr(names(kesFlow)[a + 1], 4,6))
 dfnames[a] = paste0("Alt", substr(names(kesFlow)[a + 1], 4,6))
 names(dflist) = dfnames # naming the list 
 
   #table values
  avg_sept_flow <- round(mean(as.numeric(septKesFlow[, paste0("Alt", substr(names(kesFlow)[a + 1], 4,6))]), na.rm = TRUE), 0)
  
  avg_oct_flow <- round(mean(as.numeric(octKesFlow[, paste0("Alt", substr(names(kesFlow)[a + 1], 4,6))]), na.rm = TRUE), 0) 
  total_volume_sf <- round((sum(as.numeric(kesFlow[kesFlow$Date >= "2023-09-01" & kesFlow$Date < "2024-03-01" , names(kesFlow)[a + 1]]))*1.983)/1000, 0)
  total_volume_ao <- round((sum(as.numeric(kesFlow[kesFlow$Date >= "2023-08-01" & kesFlow$Date < "2023-11-01" , names(kesFlow)[a + 1]]))*1.983)/1000, 0)
  
  wr_dewatered <- sum(redds[ ,paste0("dewateredAlt", substr(names(kesFlow)[a + 1], 4,6))], na.rm = TRUE)
  wr_pct_lost <-  round((sum(redds[ ,  paste0("dewateredAlt", substr(names(kesFlow)[a + 1], 4,6))], na.rm = TRUE)/reddCount*100), 1)
  
  wr_pct_lost_70 <- round((sum(redds[ ,  paste0("dewateredAlt", substr(names(kesFlow)[a + 1], 4,6))], na.rm = TRUE)/(reddCount*1.7)*100), 1)
  
  wr_pct_lost_30 <- round((sum(redds[ ,  paste0("dewateredAlt", substr(names(kesFlow)[a + 1], 4,6))], na.rm = TRUE)/(reddCount*1.3)*100), 1)
  
  wr_dewatered_250 <- sum(redds[ ,  paste0("dewateredAlt", substr(names(kesFlow)[a + 1], 4,6), "Buffer")], na.rm = TRUE)
  
  wr_pct_lost_250 <- round((sum(redds[ ,  paste0("dewateredAlt", substr(names(kesFlow)[a + 1], 4,6), "Buffer")], na.rm = TRUE)/reddCount*100), 1)
 
 dflist[[a]] <- as.character(c(avg_sept_flow, avg_oct_flow, total_volume_sf, total_volume_ao, 
                               wr_dewatered, wr_pct_lost, wr_pct_lost_70, wr_pct_lost_30,
                                wr_dewatered_250, wr_pct_lost_250))

 table <- cbind(table, dflist[a])
}



table <- rbind(table, setNames(table2, names(table)))
knitr::kable(table)
```

# Plots
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.dim = c(6,3.5)}
reddsDat <- reddsAll
reddsTable <- setDT(reddsDat)[,list(Count=.N), names(reddsDat)]
ymin <- round_any(min(reddsTable$ACTUAL.or.ESTIMATED..DEWATER.FLOW), 1000, f = floor)
#ymax <- round_any(max(reddsTable$ACTUAL.or.ESTIMATED..DEWATER.FLOW), 250, f = ceiling)
ymax <- round_any(max(realFlow$KES_Flow), 250, f = ceiling)
todays_date <- as.Date(format(max(kwk_flows$Date), '%Y-%m-%d'))
mid <- as.Date('2023-09-01') + floor((todays_date - as.Date('2023-09-01'))/2)
for(r in 1:nrow(reddsTable)){
  if(reddsTable$Status[r] == "EMERGED"){
    reddsTable$cols[r] <- "Emerged"
  } else {
    if(reddsTable$Status[r] == "DEWATERED"){
      reddsTable$cols[r] <- "Dewatered"
    } else {
      reddsTable$cols[r] <- "Remaining"
    }
  }
}
reddsTable <- reddsTable %>% rename('Date' = 'Estimated.Date.of.Emergence', 'Flow' = 'ACTUAL.or.ESTIMATED..DEWATER.FLOW') 

flows <- kesFlowReal %>% 
  gather(key = Alts, value = Flow, contains('cfs')) %>%
  mutate(Alts = str_extract(Alts, "\\d([a-zA-Z])")) %>% 
  mutate(Flow = as.numeric(Flow)) %>%
  filter(Date >= todays_date)

real_flows <- kes_flows %>% left_join(kwk_flows, by = 'Date') %>%
  gather(key = 'Gauge', value = 'Flow', KES, KWK) %>% mutate(Flow = as.numeric(Flow))

ggplot() + geom_line(flows, mapping = aes(x = Date, y = Flow, color = Alts), size = 0.75) +
  geom_line(real_flows, mapping = aes(x = Date, y = Flow, linetype = Gauge), size = 0.75) +
  scale_linetype_discrete(guide = "none") +
  geom_label(reddsTable, mapping = aes(x = Date, y = Flow, 
                                       label = Count, fill = cols), min.segment.length = 4,
             postion = 'dodge', size = 2.5, label.padding = unit(0.25, 'lines'),label.r = unit(.35, 'lines'), fontface = 'bold', alpha = 0.75) + #facet_wrap(~Alternates) +
  ylim(ymin, ymax) +
  xlim(min(as.Date('2023-09-01')), (max(reddsTable$Date) + 5)) +
  labs(x = 'Date', y = 'Flow (cfs)', fill = 'Redd Status') +
  theme(legend.position = 'bottom', 
        legend.text = element_text(size = 8), 
        axis.text = element_text(size = 8),
        axis.title = element_text(size = 10),
        legend.title = element_text(size = 9)) +
  scale_fill_manual(labels = c('Remaining', 'Emerged', 'Dewatered'), values = c('grey', 'lightblue', 'darkorange')) +
  annotate(geom = 'rect', xmin = as.Date('2023-09-01'), xmax = max(todays_date), 
           ymin = 4000, ymax = 10000, fill = 'darkgrey', color = 'black', alpha = 0.3, linetype = 'dotted') +
  annotate(geom = 'text', x = (mid + 1), y = 10300, size = 3, fontface = 'italic', label = 'Real-time Flows') +
  annotate(geom = 'text', x = as.Date('2023-09-03'), y = 9350, size = 2.25, label = 'KES') +
  annotate(geom = 'text', x = as.Date('2023-09-03'), y = 7750, size = 2.25, label = 'KWK')



#text(x=rep(max(kesFlow$Date)+3, 5), 
 #    y=c(min(kesFlow$Alt1), min(kesFlow$Alt2), 
  #       min(kesFlow$Alt3), min(kesFlow$Alt4), min(kesFlow$Alt5)), 
   #  pos=4, labels=c('red line', 'yellow line', 'orange line', 'green line', 'blue line'))
#data.frame(date = unique(redds$Estimated.Date.of.Emergence), )

```

**Figure 1.** Actual or estimated emergence dates of SRWC redds and actual or estimated dewatering flow for the September-October estimated redd emergence dates as compared to Keswick flow (in cfs) of proposed management alternatives. Points represent emerged, dewatered, or remaining redds. Numbers inside of points indicate how many redds share that estimated emergence date and actual/estimated dewatering flow. Points that fall above/to the right of a flow alternative line are expected to be dewatered given that management alternative is followed. Points that fall below/to the left of/on a flow alternative line are not expected to be dewatered, given that management alternative is followed. Black line equals KWK gauge flow and navy blue line equals KES flow (from [SacPas](https://www.cbr.washington.edu/sacramento/data/query_river_table.html)). Dotted vertical gray line equals today's date.


# References
Gard, Mark. 2006. Relationships between flow fluctuations and redd dewatering and juvenile stranding for Chinook Salmon and Steelhead in the Sacramento River between Kesewick Dam and Battle Creek. 94 pages.

