---
title: 'PRELIMINARY DATA: Redd Dewatering Estimates for Keswick Fall Flow Scenarios'
author: "BDO Science Division"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document: default
  word_document: default
  pdf_document: default
---

```{r setup, echo=FALSE, include = FALSE}
#loading libraries and pulling in relevant data from excel files
library(tinytex)
library(data.table)
library(plyr)
library(ggplot2)
library(tidyverse)
library(readxl)
library(zoo)
library(kableExtra)
library(janitor)
library(lubridate)

#set desired scenarios for later filtering
#scenarios <- c("Alt.1c", "Alt.2e", "Alt.3m", "Alt.3r", "Alt.3s") 

#this function will convert the character values (from xlsx format of numbers with commas) to numeric class values
setClass("num.with.commas")
setAs("character", "num.with.commas", 
        function(from) as.numeric(gsub(",", "", from) ) )

#setting yr for automating certain lables and filtering
yr = year(Sys.Date())
flow_date <- Sys.Date() #for later filterning
yr_exp <- 2 #this year's expected expansion number based on the linear relationship between yearly expansions vs recapture rate of tagged female salmon

########Read in redd data from shallow winter redd monitoring
redd_files <- list.files('External_data/ShallowRedds/', pattern = "\\.xlsx$", full.names = TRUE) #list all excel files
MaxReddFile <- max(redd_files) #reads the latest redd excel file
reddsheets <- excel_sheets(MaxReddFile) #list all sheets
reddsheet <- reddsheets[reddsheets == 'SHALLOW REDDS'] #filter sheet with shallow redd nfo

redds <- read_excel(MaxReddFile,  sheet = reddsheet, #read in shallow redd file and clean up.
                    range = cell_cols(c('A:I'))) %>% 
  na.omit() %>% 
  select(3,4,6,9) %>%
  rename('Born.on.Date' = 1, 'Estimated.Date.of.Emergence' = 2, 'ACTUAL.or.ESTIMATED..DEWATER.FLOW' = 4) %>%
  mutate_at(1:2, as.Date)#minor cleaning

reddsAll <- redds

########Read in most recent flow scenario data, cleaning up datasheet, etc.
Files <- list.files('External_data/FlowScen/', pattern = "xlsx$")
MaxFile <- max(Files)
flowsheets <- excel_sheets(paste0('External_data/FlowScen/',MaxFile))
flowsheet <- grep("Alternatives", flowsheets, value = TRUE) #for pulling in flow alternatives
scensheet <- grep("Desired", flowsheets, value = TRUE) #for pulling in desired scenarios sheet

#pull in desired scenarios
scen <- read_excel(paste0('External_data/FlowScen/',MaxFile), 
                      sheet = scensheet, col_names = TRUE) %>%
  filter(Use == 'Y') %>% 
  mutate(Scenario = gsub("[^[:alnum:]]+", "", Scenario)) %>%
  mutate(Scenario = tolower(Scenario)) %>%
  pull(Scenario)

#pull in flow alternatives
kesFlow <- read_excel(paste0('External_data/FlowScen/',MaxFile), 
                      sheet = flowsheet, skip = 1, col_names = TRUE) 
colnames(kesFlow) <- gsub("[^[:alnum:]]+", "", colnames(kesFlow))
kesFlow <- kesFlow %>%
  clean_names() %>%
  mutate(date = as.Date(as.numeric(date), origin = "1899-12-30")) %>%
  filter(!is.na(date)) %>%
  select(-contains("actual"), -contains("timeline")) %>%  # Exclude columns with 'Actual' and 'Timeline'
  select_if(~ !all(is.na(.))) %>% #Exlcude blank columns
  gather(key = scenarios, value = flow, -date) %>% #gather for easier filtering
  mutate(scenarios = gsub('_', '', scenarios)) %>%
  filter(scenarios %in% scen) %>% #filter for desired scenarios
  spread(key = scenarios, value = flow) #spread back out for later use

########Import relevant fall-run redd dewatering data

#import fall-run spawn info
spawn <- read.csv('Model_inputs/spawn_timing.csv') %>% 
  mutate(Date = as.Date(Day, origin = as.Date(paste0(yr,'-01-01')))) %>% #converts Julian Day to date for current year
  rename('EmergDays' = 'EmergDate') %>% #renaming for later use
  filter(Run == 'Fall') #filtering just for fall-run

#import gard look up table
fall_lookup <- read.csv('Model_inputs/model_dewater_flows.csv') %>%
  rename('GardDewater' = 'Dewater', 'GardSpawn' = 'Spawn_Flows') %>% 
  filter(Run == 'Fall')

########read in Redd Count data and date from most recent file

count_pattern <- 'To date, unexpanded redd count' #set pattern for count cell to look for
date_pattern <- 'Through' #set pattern for date cell to look for
CountFiles <- list.files('External_data/ShallowRedds/ReddCount/', 
                         pattern = "xlsx$") #list files with reporting tab in them
MaxCountFile <- max(CountFiles) #single out the most recent file
sheetCount <- data.frame(excel_sheets(paste0('External_data/ShallowRedds/ReddCount/', 
                                             MaxCountFile))) %>% #read in most recent excel WITH count data
  rename('name' = 1) %>% 
  filter(grepl('REPORTING', name, ignore.case = TRUE)) #lists all excel sheets in the file

#for Count data
Count <- read_excel(paste0('External_data/ShallowRedds/ReddCount/',MaxCountFile),  
                    sheet = sheetCount[1,1]) #pull in sheet with count data
reddCount <- round(as.numeric(Count[[(which(Count[, 2] == 'To date, unexpanded redd count') + 1), 2]]),0) #isolate count
countDate <- format(as.Date(as.numeric(Count[[(which(Count[, 1] == 'Through') + 1), 1]]), 
                            origin = "1899-12-30"), "%B %d, %Y") #isolate count data
updatedReddInfoDate <- format(as.Date(gsub(".*?(\\d{4}-\\d{2}-\\d{2}).*", "\\1", MaxReddFile), format = "%Y-%m-%d"), "%B %d, %Y") #isolate most recent redd info date from file name

```

This script constructs real-time winter-run redd dewatering estimates based on most recent data available from CDFW (`r updatedReddInfoDate`) for winter-run data and dewatering estimates from USFWS (2006; see citation). Data are also available in `r yr` Winter-run Data file.xls online at [calfish.org](https://gcc02.safelinks.protection.outlook.com/?url=https%3A%2F%2Fwww.calfish.org%2FProgramsData%2FConservationandManagement%2FCentralValleyMonitoring%2FCDFWUpperSacRiverBasinSalmonidMonitoring.aspx&data=05%7C01%7Clelliott%40usbr.gov%7C689ebb9a6c8243b4f96c08da90f5c542%7C0693b5ba4b184d7b9341f32f400a5494%7C0%7C0%7C637981682646098788%7CUnknown%7CTWFpbGZsb3d8eyJWIjoiMC4wLjAwMDAiLCJQIjoiV2luMzIiLCJBTiI6Ik1haWwiLCJXVCI6Mn0%3D%7C3000%7C%7C%7C&sdata=A1eQkWPxbkXxnzEvc2K8%2FTmslZ8H8zvxdks3%2F78Yrvw%3D&reserved=0).

Please note that all data are preliminary until data collection is finalized. Likewise, there are uncertainties with forecasts which may lead to changes in proposed operations.

```{r, echo = FALSE, include = FALSE}
#####pulling in real-time flow data for KES and KWK

# query parameters for SacPAS
# calendar year
queryYear=yr

# KWK river flow data from SacPAS
queryKWKflow= paste0("https://www.cbr.washington.edu/sacramento/data/php/rpt/mg.php?sc=1&mgconfig=river&outputFormat=csvSingle&hafilter=All&year%5B%5D=",yr,"&loc%5B%5D=KWK&data%5B%5D=Flow&tempUnit=F&startdate=1%2F1&enddate=12%2F31&avgyear=0&consolidate=1&grid=1&y1min=&y1max=&y2min=&y2max=&size=large")
kwkFlowData = read.csv(queryKWKflow) %>% slice(1:(n()-3))

# KES reservoir flow data from SacPAS
queryKESflow=paste0("https://www.cbr.washington.edu/sacramento/data/php/rpt/mg.php?sc=1&mgconfig=river&outputFormat=csvSingle&hafilter=All&year%5B%5D=",yr,"&loc%5B%5D=KES&data%5B%5D=ReservoirOutflow&tempUnit=F&startdate=1%2F1&enddate=12%2F31&avgyear=0&consolidate=1&grid=1&y1min=&y1max=&y2min=&y2max=&size=large")
kesFlowData = read.csv(queryKESflow) %>% slice(1:(n()-3))

allflows <- rbind(kwkFlowData, kesFlowData)
allflows <- allflows %>% 
  mutate(Date = as.Date(paste0(mm.dd,'-',year), format = '%m-%d-%Y')) %>% 
  rename('Flow' = 'value') %>% 
  select(Date, Gage = location, Flow) %>%
  filter(Date >= '2024-08-01' & Date <= Sys.Date()-1)
```

```{r, echo = FALSE}
# Converting flows to Thousand Acre Feet and overlaying real time flows over flow scenarios as season progresses
kes <- allflows %>% filter(Gage =='KES') %>% select(-Gage, Date, KES = 'Flow')
names <- colnames(kesFlow)
names <- names[names != "date"]
kesTemp <- data.frame(Date = kes$Date)

for(i in names) {
  name <- i
  temp <- data.frame(name = kes$KES)
  kesTemp <- cbind(kesTemp, temp)
}

colnames(kesTemp) <- colnames(kesFlow)

kesFlowReal <- bind_rows(kesTemp, filter(kesFlow, date > max(kes$Date))) 


kesFlowReal <- kesFlowReal %>% 
  rename_with(~ paste0(., "_cfs"), -date) %>%  # Rename original columns with '_cfs'
  mutate(across(ends_with("_cfs"), ~ . * 1.983 / 1000, .names = "{.col %>% str_replace('_cfs$', '')}_taf"))

septTAF <- subset(kesFlowReal, months(kesFlowReal$date) == "September")
octTAF <- subset(kesFlowReal, months(kesFlowReal$date) == "October")

```

```{r, echo = FALSE}
# September Flow Average
septKesFlow <- subset (kesFlowReal, kesFlowReal$date < paste0(yr,'-10-01') & kesFlowReal$date >= paste0(yr, '-9-01'))
septKesFlow <- septKesFlow %>% select(date, contains("_cfs"))
octKesFlow <- subset (kesFlowReal, kesFlowReal$date >= paste0(yr,'-10-01') & kesFlowReal$date < paste0(yr, '-11-01'))
octKesFlow <- octKesFlow %>% select(date, contains("_cfs"))
```

```{r, echo = FALSE, message=FALSE}
#estimate fall-run redds dewatered
#function for converting flows to nearest number in Gard lookup
round_to_nearest <- function(x, basis_vector) {
  nearest_value <- basis_vector[which.min(abs(basis_vector - x))]
  return(nearest_value)
}
#transform kesFlowReal dataframe for model
redd_model <- kesFlowReal %>% 
  gather(key = Scenarios, value = Spawn_Flows, contains('cfs')) %>%
  select(date, Scenarios, Spawn_Flows) %>% 
  mutate(Boards = if_else(date < as.Date(paste0(yr,'-11-01')), 'in', 'out')) %>% 
  mutate(Spawn_Flows = as.numeric(Spawn_Flows)) %>% 
  left_join(spawn, by = c('date'= 'Date')) %>% 
  filter(!is.na(Run)) %>%
  mutate(Spawn_Flows = na.approx(Spawn_Flows), EmergDate = date + EmergDays) %>%
  mutate(Dewater = sapply(EmergDate, function(edate) {
    if (!is.na(edate)) {
      min(Spawn_Flows[date <= edate & date >= date], na.rm = TRUE)
    } else {
      NA
    }
  })) %>%
  mutate(Boards = if_else(month(date) < 11, 'in', 'out')) 
redd_model <- redd_model %>%
  mutate(GardSpawn = sapply(redd_model$Spawn_Flows, round_to_nearest, fall_lookup$GardSpawn)) %>%
  mutate(GardDewater = sapply(redd_model$Dewater, round_to_nearest, fall_lookup$GardDewater)) %>%
  left_join(fall_lookup, by = c('GardDewater', 'GardSpawn', 'Boards')) %>%
  mutate(Prop_dewater = if_else(is.na(Prop_dewater), 0, Prop_dewater)) %>%
  mutate(TotalDewater = Prop * Prop_dewater)

summary <- redd_model %>% group_by(Scenarios) %>% 
  summarize(Redds_lost = round((sum(TotalDewater)*100), 1)) %>%
  rename('Fall-run dewatered (%)' = 'Redds_lost')
table2 <- setDT(as.data.frame(t(summary)), keep.rownames = TRUE, check.names = TRUE)
names(table2) <- table2 %>% slice(1) %>% unlist()
table2 <- table2 %>% 
  slice(-1) %>%
  rename('Metric' = 1)
```

# Current Winter-run Chinook Salmon Redd Count

As of `r countDate`, the unexpanded redd count is **`r reddCount`** Winter-run redds. It is important to note that until data collection is completed for the year these are the **minimum** number of possible redds. The Winter-run number will always expand upon final analysis but gives an in-season guard rail of the minimum number of redds this year.

Given that the number of Winter-run redds is always larger than the early season carcass counts, an expansion number based on historic data is multiplied by the carcass count to estimate the total number of redds for the season before the end of the season's final estimate is developed and the final redd count is known. These additional estimates of redd counts (shown in Table 1) help to inform decisions regarding possible redd dewatering. However, expansion estimates have not been developed yet for the 2024 season and those in Table 1 are from the 2023 season.

```{r, echo = FALSE, message=FALSE}
# create a table for population expansions
pop_tab <- data.frame(Name = c("Current Count", paste0("Expected ", yr, " Expansion"), "Maximum Expansion", "Minimum Expansion"),
                      Expansion = c(1, yr_exp, 3.45, 1.25)
                      
)

pop_tab <- pop_tab %>%
  mutate(Redds = round(reddCount*Expansion, 0),
    "1%" = Redds*0.01) %>%
  rename("Expansion Number" = Expansion,
    "Total Redds" = Redds)

knitr::kable(pop_tab,
             caption = 'Estimated total number of Winter-run redds and resulting number of redds that represent 1% of the population. Estimated total redds are based on current count and expansion numbers representing 1) average 2005-2022 expansion, 2) year-specific expansion determined by the linear relationship between yearly expansions vs recapture rate of tagged female salmon, 3) maximum 2005-2022 expansion, and 4) minimum 2005-2022 expansion.')
```

## Chinook Salmon Dewatered Redd Estimates

As of `r updatedReddInfoDate`, **`r nrow(subset(reddsAll, reddsAll$Status == "EMERGED"))`** Winter-run redds have **emerged** and **`r nrow(subset(reddsAll, reddsAll$Status == "DEWATERED"))`** have been **dewatered**. This leaves **`r nrow(subset(reddsAll, reddsAll$Status == "OK"))`** shallow water redds of concern.

There is no real time data on fall-run redd counts. Estimates are predicted based on estimated dewatering percentages from USFWS (2006) and spring-run and fall-run spawn timing based on fresh female carcasses encountered by week from 2003 through 2022. Emergence timing were predicted from average water temperatures over the same time period. Fall-run dewatered redd estimates range from **`r min(summary[2])`** to **`r max(summary[2])`%**.  Note that fall-run dewatering estimates are likely overestimated using the dewatering percentages from USFWS (2006), and likely do not reflect actual dewatering percentages and should only be used for comparative purposes between scenarios.  A comparative analysis between field and modeled dewatering percentages by Gosselin and Beer (2024) can be found here: https://www.cbr.washington.edu/sacramento/fishmodel/Note_on_Redd_Dewatering_Observed_v_Predicted.pdf.

```{r, echo = FALSE}
library(lubridate)

# Estimated number of redds dewatered
kesFlowReal2 <- kesFlowReal %>% select(date, contains("_cfs"))
kesFlowReal3 <- kesFlowReal2 %>% gather(key = 'Alts', value = 'Flow', -date)
name2 <- colnames(kesFlowReal2)
name2 <- name2[name2 != "date"]

#formatting new redds file for figuring out minimum flows
redds2 <- subset(redds, redds$Status == "OK" | redds$Status == "DEWATERED")

for (col in name2) { #adding alternative names to redd dataframe
  redds2[[col]] <- NA
}

redds2 <- redds2 %>% gather(key = 'Alts', value = 'Flow', -1:-4) %>%
  select(-Flow)
flowList <- list()

for(i in 1:nrow(redds2)){
  filtered <- kesFlowReal3 %>% filter(Alts == redds2$Alts[i]) %>%
    filter(date >= redds2$Born.on.Date[i] & date <= redds2$Estimated.Date.of.Emergence[i])
  minflow <- min(filtered$Flow)
  flowList[[i]] <- minflow
}

test2 <- cbind(flowList)
dewater <- cbind(redds2, test2)
dewater <- dewater %>% mutate(Dewater = if_else(flowList <= ACTUAL.or.ESTIMATED..DEWATER.FLOW, 1, 0)) %>%
  mutate(Dewater_Buffer = ACTUAL.or.ESTIMATED..DEWATER.FLOW + 250) %>%
  mutate(Dewatered_Buffer = if_else(flowList <= Dewater_Buffer, 1, 0)) %>%
  gather(key = 'Type', value = 'Dewater', 7,9) %>%
  mutate(Buffer = if_else(grepl('Buffer', Type), 'YES', 'NO')) %>%
  select(5,9,10) %>%
  mutate(key = row_number())
  
dewater <- dewater %>% spread(key = 'Alts', value = 'Dewater')

```



```{r, echo=FALSE, message = FALSE}
#automating expansion row names for table
exp <- vector()
pop_temp <- pop_tab %>% rename('Expansion' = 'Expansion Number') %>% filter(Expansion > 1, !grepl('Expected', Name))

for(i in 1:nrow(pop_temp)){
  temp <- pop_temp[i,'Expansion']
  temp <- paste0('Winter-run Percent Lost (expansion of ', temp,')')
  exp[i] <- temp
}
#creating row names for table
Metric <- c("Avg Sept Flow (cfs)", "Avg Oct Flow (cfs)", "Sept-Feb Total Volume (TAF)", "Aug-Oct Total Volume (TAF)", "Winter-run Redds Dewatered", "Winter-run Percent Lost (current count)", paste0("Winter-run Percent Lost (expansion of ", yr_exp, ")"),exp , "Winter-run Redds Dewatered (250 cfs buffer)", "Winter-run Percent Lost (250 cfs buffer)")

#creating the table using the row names
table <- data.frame(Metric)

#creating a list to store all summary statistics below
tableList <- list()

#summary statistics for flow which are stored in tableList
tableList <- append(tableList, list(avg_sept_flow = colMeans(septKesFlow[,-1], na.rm = TRUE)), after = length(tableList))
tableList <- append(tableList, list(avg_oct_flow = colMeans(octKesFlow[,-1], na.rm = TRUE)), after = length(tableList))
tableList <- tableList %>% append(
    list(total_volume_sf = (colSums(filter(kesFlowReal2, date >= paste0(yr,'-09-01') & date < paste0(yr+1,'-03-01'))[, -1])*1.983/1000)),
    after = length(tableList)
  )
tableList <- tableList %>% append(
    list(total_volume_ao = (colSums(filter(kesFlowReal2, date >= paste0(yr,'-08-01') & date < paste0(yr,'-11-01'))[, -1])*1.983/1000)),
    after = length(tableList)
  )

#summary statistics for redd dewatering
tableList <- tableList %>% append(
    list(wr_dewatered = (colSums(filter(dewater, Buffer == 'NO')[, -1:-2], na.rm = TRUE))),
    after = length(tableList)
  )
tableList <- tableList %>% append(
    list(wr_pct_lost = (colSums(filter(dewater, Buffer == 'NO')[, -1:-2], na.rm = TRUE)/reddCount)*100),
    after = length(tableList)
  )

#summary statistics for redd dewatering with expansion factors
tableList <- tableList %>% append(
    list(wr_pct_lost_300 = (colSums(filter(dewater, Buffer == 'NO')[, -1:-2], na.rm = TRUE)/(reddCount*3))*100),
    after = length(tableList)
  )
tableList <- tableList %>% append(
    list(wr_pct_lost_345 = (colSums(filter(dewater, Buffer == 'NO')[, -1:-2], na.rm = TRUE)/(reddCount*3.45))*100),
    after = length(tableList)
  )

#summary statsitics for redd dewatering with flow buffer
tableList <- tableList %>% append(
    list(wr_pct_lost_125 = (colSums(filter(dewater, Buffer == 'NO')[, -1:-2], na.rm = TRUE)/(reddCount*1.25))*100),
    after = length(tableList)
  )
tableList <- tableList %>% append(
    list(wr_dewatered_250 = (colSums(filter(dewater, Buffer == 'YES')[, -1:-2], na.rm = TRUE))),
    after = length(tableList)
  )
tableList <- tableList %>% append(
    list(wr_pct_lost_250 = (colSums(filter(dewater, Buffer == 'YES')[, -1:-2], na.rm = TRUE)/reddCount)*100),
    after = length(tableList)
  )

#binding summary statistics into one table
temp_table <- bind_rows(tableList)

#binding to original table with row names
table <- bind_cols(Metric, temp_table) %>% rename('Metric' = 1)

#adding fall-run dewatering to table
table <- rbind(table, setNames(table2, names(table))) %>% 
  mutate_at(c(2:ncol(table)), as.numeric)
# Separate the data into two parts: the first 5 rows and the remaining rows
table[,2:ncol(table)] <- round(table[,2:ncol(table)], 2)
# Print the modified table
colnames(table) <- gsub("_cfs$", "", colnames(table))

knitr::kable(table,
             caption = paste0('Average September and October Keswick (KES) Flow in cfs, total water volume of each alternative for August through October and September through February in TAF, estimated numbers of SRWC redds dewatered, and percent of population that would be lost under each of the proposed alternatives. KES Flow data uses actual flow-to-date as of ',format(max(allflows$Date), '%B %d, %Y'),' and proposed flows for the remainder of the incubation period. Redd dewatering is considered at the actual or estimated dewatering flow and with a 250 cfs buffer applied to the actual/estimated dewatering flow. Percentage of the population lost is based on the ',countDate,' count of ',reddCount,' Winter-run redds and updated redd counts may be available soon. See Scenario Descriptions file for additional information on each scenario.'))
```



```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.cap = 'Actual or estimated emergence dates of SRWC redds and actual or estimated dewatering flow for the September-October estimated redd emergence dates as compared to Keswick flow (in cfs) of proposed management alternatives. Points represent dewatered (De), emerged (Em), or remaining (Re) redds. Numbers inside of points indicate how many redds share that estimated emergence date and actual/estimated dewatering flow. Points that fall above/to the right of a flow alternative line are expected to be dewatered given that management alternative is followed. Points that fall below/to the left of/on a flow alternative line are not expected to be dewatered, given that management alternative is followed. Shaded gray box shows period of real-time flow data; dashed black line equals KWK gauge flow and solid black line equals KES flow (from [SacPas](https://www.cbr.washington.edu/sacramento/data/query_river_table.html)).'}

#, fig.dim = c(8,5)
reddsDat <- reddsAll
reddsTable <- setDT(reddsDat)[,list(Count=.N), names(reddsDat)]
ymin <- round_any(min(reddsTable$ACTUAL.or.ESTIMATED..DEWATER.FLOW), 1000, f = floor)
#ymax <- round_any(max(reddsTable$ACTUAL.or.ESTIMATED..DEWATER.FLOW), 250, f = ceiling)
ymax <- round_any(max(allflows$Flow)+250, 500, f = ceiling)
todays_date <- Sys.Date()
mid <- as.Date(paste0(yr,'-08-01')) + floor((todays_date - as.Date(paste0(yr,'-08-01')))/2)
for(r in 1:nrow(reddsTable)){
  if(reddsTable$Status[r] == "EMERGED"){
    reddsTable$cols[r] <- "Emerged"
  } else {
    if(reddsTable$Status[r] == "DEWATERED"){
      reddsTable$cols[r] <- "Dewatered"
    } else {
      reddsTable$cols[r] <- "Remaining"
    }
  }
}
reddsTable <- reddsTable %>% 
  rename('Date' = 'Estimated.Date.of.Emergence', 'Flow' = 'ACTUAL.or.ESTIMATED..DEWATER.FLOW') %>% 
  mutate(cols = factor(cols, levels = c('Remaining', 'Emerged', 'Dewatered'), labels = c('Re','Em', 'De')))

flows <- kesFlowReal %>% 
  gather(key = Alts, value = Flow, contains('cfs')) %>%
  filter(date >= todays_date) %>%
  mutate(Alts = gsub('_cfs','',Alts))

ggplot() + geom_line(flows, mapping = aes(x = date, y = Flow, color = Alts), size = 0.75) +
  geom_line(allflows, mapping = aes(x = Date, y = Flow, linetype = Gage), size = 0.75) +
  scale_linetype_discrete(guide = "none") +
  geom_label(reddsTable, mapping = aes(x = Date, y = Flow, 
                                       label = Count, fill = cols), min.segment.length = 4,
             postion = 'dodge', size = 3, label.padding = unit(0.25, 'lines'),label.r = unit(.35, 'lines'), 
             fontface = 'bold', alpha = 0.6) +
  ylim(ymin, ymax) +
  xlim(min(as.Date(paste0(yr,'-08-01'))), (max(reddsTable$Date) + 5)) +
  labs(x = 'Date', y = 'Flow (cfs)', fill = 'Redd Status') +
  theme(legend.position = 'bottom', 
        legend.text = element_text(size = 12), 
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 12),
        legend.title = element_text(size = 14),
        legend.key.size = unit(0.3, 'cm')) +
  scale_fill_manual(values = c('lightgrey', 'steelblue3', 'darkorange')) +
  annotate(geom = 'rect', xmin = as.Date(paste0(yr,'-08-01')), xmax = max(todays_date), 
           ymin = 3000, ymax = max(allflows$Flow) + 100, fill = 'darkgrey', color = 'black', alpha = 0.2, linetype = 'dotted') +
  annotate(geom = 'text', x = (mid + 0.5), y = max(allflows$Flow) + 400, size = 3.5, 
           fontface = 'italic', label = 'Real-time Flows') +
  annotate(geom = 'text', x = as.Date('2023-09-03'), y = 9350, size = 3, label = 'KWK') +
  annotate(geom = 'text', x = as.Date('2023-09-03'), y = 7750, size = 3, label = 'KES')

```


```{r echo=FALSE, message=FALSE}
worksheetSD <- read_xlsx(paste0('External_data/FlowScen/',MaxFile), sheet = "Scenario Description",range="A1:A1000",col_names=F,col_types="text")
# remove extraneous lines, keeping lines with "=" in it
keep=apply(worksheetSD,2,function(x.in) str_detect(x.in,"="))
worksheetSD=worksheetSD[keep,]
worksheetSD=worksheetSD[!is.na(worksheetSD[,1]),]
# split the Alts from the descriptions
num.in = dim(worksheetSD)[1]
scenarioDesc=data.frame(matrix(nrow=num.in,ncol=2))
colnames(scenarioDesc)=c("Scenario","Descripion")
# loop through dataframe, split original dataframe column 1 on literal "=" characters and trim leading and trailing space, create new dataframe with 2 columns "Scenario" and "Description"
# populates scenarioDesc dataframe
for(i in 1:num.in){
	scenarioDesc[i,]=str_trim(str_split(worksheetSD[i,1],"=")[[1]])
}

scenarioDesc <- scenarioDesc %>% 
  mutate(Scenario = gsub("[^[:alnum:]]+", "", Scenario)) %>%
  mutate(Scenario = tolower(Scenario)) %>%
  filter(Scenario %in% scen)
#use scenarios defined in first chunk of code to filter the scenarioDesc dataframe



#filter based on scenario name
#scenarioDesc <- scenarioDesc %>% 
  #filter(grepl(paste(toupper(scenarios2), collapse ='|'), toupper(Scenario)))
knitr::kable(scenarioDesc, caption = 'Description of scenarios being considered and compared by the Upper Sacramento Scheduling Team.') %>%
  column_spec(1, width = "2cm") %>%
  column_spec(2, width = "14cm")
```
# References

Gard, Mark. 2006. Relationships between flow fluctuations and redd dewatering and juvenile stranding for Chinook Salmon and Steelhead in the Sacramento River between Kesewick Dam and Battle Creek. 94 pages.

Gosselin, J.L. and W.N. Beer. 2024. Sacramento River Winter-run Chinook Salmon Redd Dewatering: a Note on Comparing Observed and Predicted. Central Valley Prediction and Assessment of Salmon (SacPas; https://www.cbr.washington.edu/sacramento/). Columbia Basin Research, School of Aquatic and Fishery Sciences, University of Washington.
