---
title: "PRELIMINARY DATA: Redd Dewatering Estimates for Keswick Fall Flow Scenarios"
author: "BDO Science Division"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r, echo=FALSE}
reddCount <- 347
countDate <- "August 13, 2023"
updatedReddInfoDate <- "August 31, 2023"
fallreddCount<- 1000

scenarios <- c("Alt.1b", "Alt.2d", "Alt.3d", "Alt.3f")
#show_extra = TRUE
```

This script constructs real-time winter-run redd dewatering estimates based on most recent data available from CDFW (`r updatedReddInfoDate`) for winter-run data and dewatering estimates from USFWS (2006; see citation). Data are also available in the YYYY Winter-run Data file.xls online at [calfish.org](https://gcc02.safelinks.protection.outlook.com/?url=https%3A%2F%2Fwww.calfish.org%2FProgramsData%2FConservationandManagement%2FCentralValleyMonitoring%2FCDFWUpperSacRiverBasinSalmonidMonitoring.aspx&data=05%7C01%7Clelliott%40usbr.gov%7C689ebb9a6c8243b4f96c08da90f5c542%7C0693b5ba4b184d7b9341f32f400a5494%7C0%7C0%7C637981682646098788%7CUnknown%7CTWFpbGZsb3d8eyJWIjoiMC4wLjAwMDAiLCJQIjoiV2luMzIiLCJBTiI6Ik1haWwiLCJXVCI6Mn0%3D%7C3000%7C%7C%7C&sdata=A1eQkWPxbkXxnzEvc2K8%2FTmslZ8H8zvxdks3%2F78Yrvw%3D&reserved=0).


Please note that all data are preliminary until data collection is finalized. Likewise, there are uncertainties with forecasts which may lead to changes in proposed operations. 

```{r setup, echo=FALSE, include = FALSE}
library(tinytex)
library(data.table)
library(plyr)
library(lubridate)
library(ggplot2)
library(ggrepel)
library(dplyr)
library(tidyverse)

#this function will convert the character values (from xlsx format of numbers with commas) to numeric class values
setClass("num.with.commas")
setAs("character", "num.with.commas", 
        function(from) as.numeric(gsub(",", "", from) ) )

#Read in redd data, converting flow to a numeric and date from character to date value
redds <- read.csv("Redds.csv", colClasses=c('character', 'character', 'num.with.commas'))
redds$Estimated.Date.of.Emergence <- as.Date(redds$Estimated.Date.of.Emergence, "%m/%d/%Y")
#redds$Estimated.Date.of.Emergence <- as.Date(redds$Estimated.Date.of.Emergence,  "%Y-%m-%d")
reddsAll <- redds

#Read in KES flow data, converting flow to a numeric and date from character to date value
kesFlow <- read.csv("kesFlow.csv")
kesFlow <- read.csv("kesFlow.csv", colClasses=c(rep('character', length(kesFlow))))
kesFlow$Date <- as.Date(kesFlow$Date, "%m/%d/%Y")
#kesFlow$Date <- as.Date(kesFlow$Date, "%Y-%m-%d")
kesFlow <- kesFlow %>%
  as_tibble %>%
  select(Date, scenarios) %>%
  as.data.frame()

realFlow <- read.csv("realtimeKES.csv")
realFlow$Date <- as.Date(realFlow$Date, "%m/%d/%Y")
actualDate <- max(realFlow$Date)

#import fall-run spawn info
Spawn_info <- read.csv('spawn.csv')
Spawn_info$Date <- as.Date(Spawn_info$Date, format = '%m/%d/%Y')
Spawn_info <- Spawn_info %>% mutate(Emerg_Date = as.Date(Date + Avg_emerg)) #calculate emergence date

#import fall-run model
fall_model <- readRDS('fall_model.rds')

#`r if(show_extra){
#print("this will only print when show.text is TRUE")
#str(redds)
#str(kesFlow)
#str(realFlow)}`

```

```{r, echo = FALSE}
# Thousand Acre Feet

kesFlowReal <- merge(kesFlow, realFlow, by = "Date", all.x = TRUE, all.y = TRUE)
for(a in 1:(length(kesFlow)-1)){
 kesFlowReal[ ,  paste0("A", substr(names(kesFlow)[a + 1], 4,6), "cfs")] <- NA
 kesFlowReal[ ,  paste0("A", substr(names(kesFlow)[a + 1], 4,6), "taf")] <- NA
}

for(a in 1:(length(kesFlow)-1)){
  for(r in 1: nrow(kesFlowReal)){
    if(is.na(kesFlowReal$KES_Flow[r])){
      kesFlowReal[r, paste0("A", substr(names(kesFlow)[a + 1], 4,6), "cfs")] <- kesFlowReal[r, paste0("Alt", substr(names(kesFlow)[a + 1], 4,6))]
      kesFlowReal[r, paste0("A", substr(names(kesFlow)[a + 1], 4,6), "taf")] <- (as.numeric(kesFlowReal[r, paste0("Alt", substr(names(kesFlow)[a + 1], 4,6))])*1.983)/1000
    } else{
      kesFlowReal[r, paste0("A", substr(names(kesFlow)[a + 1], 4,6), "cfs")] <- kesFlowReal$KES_Flow[r]
      kesFlowReal[r, paste0("A", substr(names(kesFlow)[a + 1], 4,6), "taf")] <- (kesFlowReal$KES_Flow[r]*1.983)/1000
    }
  }
}

septTAF <- subset(kesFlowReal, months(kesFlowReal$Date) == "September")
octTAF <- subset(kesFlowReal, months(kesFlowReal$Date) == "October")

```

```{r, echo = FALSE}
# September Flow Average
septKesFlow <- subset (kesFlow, kesFlow$Date < '2023-10-01' & kesFlow$Date >= '2023-9-01')
octKesFlow <- subset (kesFlow, kesFlow$Date >= '2023-10-01')
```

```{r, echo = FALSE, message=FALSE}
#estimate fall-run redds dewatered

#transform kesFlowReal dataframe for model
redd_model <- kesFlowReal %>% 
  gather(key = Scenarios, value = Spawn_Flows, A.1bcfs, A.2dcfs, A.3dcfs, A.3fcfs) %>%
  select(Date, Scenarios, Spawn_Flows) %>% 
  mutate(Boards = if_else(Date < as.Date('2023-11-01'), 'out', 'in')) %>% 
  mutate(Spawn_Flows = as.numeric(Spawn_Flows))

#Join the spawning information and add the minimum flow each emergence date will be subjected to
redd_model <- redd_model %>% left_join(Spawn_info, by = 'Date') %>% group_by(Scenarios) %>%
  mutate(Dewater = sapply(Emerg_Date, function(edate) {
    if (!is.na(edate)) {
      min(Spawn_Flows[Date <= edate], na.rm = TRUE)
    } else {
      NA
    }
  })) %>%
  ungroup() %>% na.omit() %>% group_by(Date) %>% mutate(Redds = fallreddCount * Per_spawn) %>% ungroup()

#apply model
pred <- predict.lm(fall_model, newdata = redd_model, type = 'response')
redd_model <- cbind(redd_model, pred) 
redd_model <- pmax(redd_model) #rounds negative predictions up to 0
redd_model <- redd_model %>% mutate(pred = if_else(Dewater >= Spawn_Flows, 0, pred)) #converts equal flows to 0

summary <- redd_model %>% mutate(Redds_lost = pred * Per_spawn) %>% group_by(Scenarios) %>% 
  summarize(Redds_lost = round((sum(Redds_lost)*100), 1)) %>%
  rename('Fall-run dewatered (%)' = 'Redds_lost')
table2 <- setDT(as.data.frame(t(summary)), keep.rownames = TRUE, check.names = TRUE)
names(table2) <- table2 %>% slice(1) %>% unlist()
table2 <- table2 %>% slice(-1)
```

# Current Winter-run Chinook Salmon Redd Count
As of `r countDate`, the unexpanded redd count is **`r reddCount`** Winter-run redds.  There is no real time data on Fall-run redd counts and instead that data is predicted based estimated dewatering percentages from USFWS (2006). It is important to note that until data collection is completed for the year these are the **minimum** number of possible redds.  The Winter-run number will always expand upon final analysis but gives an in-season guard rail of the minimum number of redds this year.  Fall-run predictions may expand or contract depending on updates from real time operations. From 2018-2021, female expansion has ranged from 0.31 to 1.31 with a 0.7 average, thus we may expect the final number of Winter-run redds to be closer to `r round(reddCount*1.7, 0)` redds using average expansion on data this year, and using this value, `r 0.01*round(reddCount*1.7, 0)` redds dewatered would be at the 1% population loss.

As of `r updatedReddInfoDate`, **`r nrow(subset(reddsAll, reddsAll$Status == "EMERGED")) `** Winter-run redds have **emerged** and **`r nrow(subset(reddsAll, reddsAll$Status == "DEWATERED"))`** have been **dewatered**. This leaves **`r nrow(subset(reddsAll, reddsAll$Status == "OK"))`** shallow water redds of concern.  Depending on real-time flow information, scenario, and 100 and 250 cfs buffers, between **`r min(summary[2]) `** and **`r max(summary[2])`%** Fall-run redds are estimated to be dewatered.



```{r, echo = FALSE}
# Estimated number of redds dewatered
for(a in 1:(length(kesFlow)-1)){
 redds[ ,  paste0("dewateredAlt", substr(names(kesFlow)[a + 1], 4,6))] <- NA
}

redds <- subset(redds, redds$Status == "OK" | redds$Status == "DEWATERED")

for(r in 1:nrow(redds)){ #loop over redds
  for(f in 1:nrow(kesFlow)){ #loop over KES flow
    for(a in 1:(length(kesFlow)-1)){ #loop over alternatives
      if(kesFlow$Date[f] == redds$Estimated.Date.of.Emergence[r]){ #if dates match
        if(kesFlow[f, paste0("Alt", substr(names(kesFlow)[a + 1], 4,6))] <= redds$ACTUAL.or.ESTIMATED..DEWATER.FLOW[r]){ #If KES flow is < or = redd dewatering flow, 1 (dewatered)
          redds[r, paste0("dewateredAlt", substr(names(kesFlow)[a + 1], 4,6))] <- 1
        } else{ #If KES flow is greater than redd dewatering flow, mark 0 (not dewatered)
          redds[r, paste0("dewateredAlt", substr(names(kesFlow)[a + 1], 4,6))] <- redds[r, paste0("dewateredAlt", substr(names(kesFlow)[a + 1], 4,6))] 
          } #ends else
      }
    } #ends if date is the same
  } #end looping over KES flow dataframe
} #end looping over redds dataframe
#Based on the `r countDate` redd count:
```


```{r, echo = FALSE}
# Estimated number of redds dewatered with a 100cfs buffer
for(a in 1:(length(kesFlow)-1)){
 redds[ ,  paste0("dewateredAlt", substr(names(kesFlow)[a + 1], 4,6), "Buffer")] <- NA
}

for(r in 1:nrow(redds)){ #loop over redds
  for(f in 1:nrow(kesFlow)){ #loop over KES flow
    for(a in 1:(length(kesFlow)-1)){ #loop over alternatives
      if(kesFlow$Date[f] == redds$Estimated.Date.of.Emergence[r]){ #if dates match
        if(as.numeric(kesFlow[f, paste0("Alt", substr(names(kesFlow)[a + 1], 4,6))]) <= redds$ACTUAL.or.ESTIMATED..DEWATER.FLOW[r] + 250){ #If KES flow is < or = redd dewatering flow, 1 (dewatered)
          redds[r, paste0("dewateredAlt", substr(names(kesFlow)[a + 1], 4,6), "Buffer")] <- 1
        } else{ #If KES flow is greater than redd dewatering flow, mark 0 (not dewatered)
          redds[r, paste0("dewateredAlt", substr(names(kesFlow)[a + 1], 4,6), "Buffer")] <- redds[r, paste0("dewateredAlt", substr(names(kesFlow)[a + 1], 4,6))] 
          } #ends else
      }
    } #ends if date is the same
  } #end looping over KES flow dataframe
} #end looping over redds dataframe
#Based on the `r countDate` redd count:

```


# Table
**Table 1.** Average September and October Keswick (KES) Flow in cfs, total water volume of each alternative for August through October in TAF, estimated numbers of SRWC redds dewatered, and percent of population that would be lost under each of the proposed alternatives. KES Flow data uses actual flow-to-date as of `r max(realFlow$Date)` and proposed flows for the remainder of the incubation period. Redd dewatering is considered at the actual or estimated dewatering flow and with a 250 cfs buffer applied to the actual/estimated dewatering flow. Percentage of the population lost is based on the `r countDate` count of `r reddCount` Winter-run redds and updated redd counts may be available soon. See Scenario Descriptions file for additional information on each scenario. 


```{r, echo=FALSE, results= TRUE}
Metric <- c("Avg Sept Flow (cfs)", "Avg Oct Flow (cfs)", "Total Volume (TAF)", "Winter-run Redds dewatered", "Winter-run Percent Lost (based on current count)", "Winter-run Percent Lost (based on expansion of 0.7)", "Winter-run Percent Lost (based on expansion of 0.3)", "Winter-run Redds Dewatered (w/ 250 cfs buffer)", "Winter-run Percent Lost (w/ 250 cfs buffer)")
table <- data.frame(Metric)


d <- rep(NA, length(kesFlow)-1)
dflist = vector("list", length(kesFlow)-1) # creating an empty list
dfnames <- rep(NA, length(kesFlow)-1)

for(a in 1:(length(kesFlow)-1)){
 d[a] = paste0("Alt", substr(names(kesFlow)[a + 1], 4,6))
 dfnames[a] = paste0("Alt", substr(names(kesFlow)[a + 1], 4,6))
 names(dflist) = dfnames # naming the list 
 
   #table values
  avg_sept_flow <- round(mean(as.numeric(septKesFlow[, paste0("Alt", substr(names(kesFlow)[a + 1], 4,6))]), na.rm = TRUE), 0)
  
  avg_oct_flow <- round(mean(as.numeric(octKesFlow[, paste0("Alt", substr(names(kesFlow)[a + 1], 4,6))]), na.rm = TRUE), 0) 
  total_volume <- round((sum(as.numeric(kesFlow[, names(kesFlow)[a + 1]]))*1.983)/1000, 0)
  
  wr_dewatered <- sum(redds[ ,paste0("dewateredAlt", substr(names(kesFlow)[a + 1], 4,6))], na.rm = TRUE)
  wr_pct_lost <-  round((sum(redds[ ,  paste0("dewateredAlt", substr(names(kesFlow)[a + 1], 4,6))], na.rm = TRUE)/reddCount*100), 1)
  
  wr_pct_lost_70 <- round((sum(redds[ ,  paste0("dewateredAlt", substr(names(kesFlow)[a + 1], 4,6))], na.rm = TRUE)/(reddCount*1.7)*100), 1)
  
  wr_pct_lost_30 <- round((sum(redds[ ,  paste0("dewateredAlt", substr(names(kesFlow)[a + 1], 4,6))], na.rm = TRUE)/(reddCount*1.3)*100), 1)
  
  wr_dewatered_250 <- sum(redds[ ,  paste0("dewateredAlt", substr(names(kesFlow)[a + 1], 4,6), "Buffer")], na.rm = TRUE)
  
  wr_pct_lost_250 <- round((sum(redds[ ,  paste0("dewateredAlt", substr(names(kesFlow)[a + 1], 4,6), "Buffer")], na.rm = TRUE)/reddCount*100), 1)
 
 dflist[[a]] <- as.character(c(avg_sept_flow, avg_oct_flow, total_volume, 
                               wr_dewatered, wr_pct_lost, wr_pct_lost_70, wr_pct_lost_30,
                                wr_dewatered_250, wr_pct_lost_250))

 table <- cbind(table, dflist[a])
}



table <- rbind(table, setNames(table2, names(table)))
knitr::kable(table)
```

# Plots
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.dim = c(6,3.5)}
reddsDat <- reddsAll
reddsTable <- setDT(reddsDat)[,list(Count=.N), names(reddsDat)]
ymin <- round_any(min(reddsTable$ACTUAL.or.ESTIMATED..DEWATER.FLOW), 1000, f = floor)
#ymax <- round_any(max(reddsTable$ACTUAL.or.ESTIMATED..DEWATER.FLOW), 250, f = ceiling)
ymax <- round_any(max(realFlow$KES_Flow), 250, f = ceiling)

for(r in 1:nrow(reddsTable)){
  if(reddsTable$Status[r] == "EMERGED"){
    reddsTable$cols[r] <- "Emerged"
  } else {
    if(reddsTable$Status[r] == "DEWATERED"){
      reddsTable$cols[r] <- "Dewatered"
  } else {
    reddsTable$cols[r] <- "Remaining"
    }
  }
}
reddsTable <- reddsTable %>% rename('Date' = 'Estimated.Date.of.Emergence', 'Flow' = 'ACTUAL.or.ESTIMATED..DEWATER.FLOW') 
flows <- kesFlowReal %>% rename('Alt-1b' = 'A.1bcfs', 'Alt-2d' = 'A.2dcfs', 
                                'Alt-3d' = 'A.3dcfs', 'Alt-3f' = 'A.3fcfs') %>%
  gather(key = 'Alternates', value = 'Flow', 'Alt-1b', 'Alt-2d', 'Alt-3d', 'Alt-3f') %>% 
  mutate(Flow = as.numeric(Flow)) %>%
  filter(Date > "2023-08-31")

ggplot() + geom_line(flows, mapping = aes(x = Date, y = Flow, color = Alternates), size = 0.75) + 
  geom_label(reddsTable, mapping = aes(x = Date, y = Flow, 
                                             label = Count, fill = cols), min.segment.length = 4,
                   postion = 'dodge', size = 2.5, label.padding = unit(0.25, 'lines'),label.r = unit(.35, 'lines'), fontface = 'bold', alpha = 0.75) + #facet_wrap(~Alternates) +
  ylim(ymin, ymax) +
  xlim(min(flows$Date), (max(reddsTable$Date) + 5)) +
  labs(x = 'Date', y = 'Flow (cfs)', fill = 'Redd Status') +
  theme(legend.position = 'bottom', 
        legend.text = element_text(size = 8), 
        axis.text = element_text(size = 8),
        axis.title = element_text(size = 10),
        legend.title = element_text(size = 10)) +
  scale_fill_manual(labels = c('Remaining', 'Emerged', 'Dewatered'), values = c('grey', 'lightblue', 'darkorange'))



#text(x=rep(max(kesFlow$Date)+3, 5), 
 #    y=c(min(kesFlow$Alt1), min(kesFlow$Alt2), 
  #       min(kesFlow$Alt3), min(kesFlow$Alt4), min(kesFlow$Alt5)), 
   #  pos=4, labels=c('red line', 'yellow line', 'orange line', 'green line', 'blue line'))
#data.frame(date = unique(redds$Estimated.Date.of.Emergence), )

```

**Figure 1.** Actual or estimated emergence dates of SRWC redds and actual or estimated dewatering flow for the September-October estimated redd emergence dates as compared to Keswick flow (in cfs) of proposed management alternatives. Points represent emerged, dewatered, or remaining redds. Numbers inside of points indicate how many redds share that estimated emergence date and actual/estimated dewatering flow. Points that fall above/to the right of a flow alternative line are expected to be dewatered given that management alternative is followed. Points that fall below/to the left of/on a flow alternative line are not expected to be dewatered, given that management alternative is followed.


# References
Gard, Mark. 2006. Relationsships between flow fluctuations and redd dewatering and juvenile stranding for Chinnok Salmon and Steelhead in the Sacramento River between Kesewick Dam and Battle Creek. 94 pages.

